Discrete Quantum Theories and Computing

### 2018-03-10 dissertation.tex
dissertation.lyx "2.2 Quantum Circuit Model" Updating...

Given a normalized pure state $\ket{\Psi}=\sum_{i=0}^{D-1}\alpha_{i}\ket{i}$,
a natural quantity characterizing an $D$-dimensional system is its
\textit{density matrix}, $\rho=\op{\Psi}{\Psi}$, or 
\begin{equation}
\rho=\begin{bmatrix}|\alpha_{0}|^{2} & \alpha_{0}\alpha_{1}^{*} & \cdots & \alpha_{0}\alpha_{D-1}^{*}\\
\alpha_{1}\alpha_{0}^{*} & |\alpha_{1}|^{2} & \cdots & \alpha_{1}\alpha_{D-1}^{*}\\
\vdots & \vdots & \ddots & \vdots\\
\alpha_{D-1}\alpha_{0}^{*} & \cdots & \alpha_{D-1}\alpha_{D-2}^{*} & |\alpha_{D-1}|^{2}
\end{bmatrix}\,.\label{nqDensity.eq}
\end{equation}
We can now use the complex generalization of the classical Veronese
coordinate system for projective geometry to remove the overall phase
ambiguity $\rme^{\rmi\theta}$ from the $D$-dimensional states. If
we take a particular weighting of the elements of the density matrix~$\rho$,
we can construct a \textit{unit vector} of real dimension $D^{2}$
with the form: 
\begin{equation}
\Hat{a}=\left(|\alpha_{i}|^{2},\ldots,\sqrt{2}\,\mathrm{Re}\ \alpha_{i}\alpha_{j}^{*},\ldots,\sqrt{2}\,\mathrm{Im}\ \alpha_{i}\alpha_{j}^{*},\ldots\right)\,,\label{nqCoords.eq}
\end{equation}
where 
\begin{equation}
\Hat{a}\cdot\Hat{a}=\sum_{i=0}^{D-1}\left(\left|\alpha_{i}\right|^{2}\right)^{2}+\sum_{i=0}^{D-1}\sum_{\substack{j=0\\
j\ne i
}
}^{D-1}\left(\mathrm{Re}\ \alpha_{i}\alpha_{j}^{*}\right)^{2}+\left(\mathrm{Im}\ \alpha_{i}\alpha_{j}^{*}\right)^{2}=\left(\sum_{i=0}^{D-1}\left|\alpha_{i}\right|^{2}\right)\left(\sum_{j=0}^{D-1}\left|\alpha_{j}\right|^{2}\right)=1\,.
\end{equation}
This construction gives an explicit embedding of the $\left(D-1\right)$-dimensional
complex, or $\left(2D-2\right)$-dimensional real, object in a real
space of dimension $D^{2}$. However, this is somewhat subtle because
the vector is of unit length, so technically the embedding space is
a sphere of dimension $D^{2}-1$ embedded in $\R^{D^{2}}$. For example,
the two-dimensional irreducible states could be represented in a four-dimensional
embedding, but the magnitude of every coordinate would be one; furthermore,
the object embedded in the resulting $\Sphere{3}$ is indeed $\Sphere{2}$
because we can fix one complex coordinate to be unity, and let one
vary, giving a total of two irreducible dimensions. In fact one must
choose \textit{two} coordinate patches, one covering one pole of $\Sphere{2}$
with coordinates $\alpha_{0}=1+\rmi0$ and $\alpha_{1}=x_{1}+\rmi y_{1}$,
and the other patch covering the other pole of $\Sphere{2}$ with
coordinates $\alpha_{0}=x_{0}+\rmi y_{0}$ and $\alpha_{1}=1+\rmi0$.

@@

\begin{comment}
An $n$-qubit pure state in the conventional quantum theory (CQT)
is represented as a vector in the $2^{n}$-dimensional Hilbert space.
If we eliminate all symmetries, an irreducible state is actually a
point in the projective Hilbert space \cite{MosseriDandoloff2001,Jaeger2007}
also known as the complex projective space~$\CP{2^{n}-1}$ \cite{Hatcher2001,Bengtsson2007}.
These irreducible quantum states can also be classified as product
states and entangled states, and the latter one plays the essential
role for pure-state algorithms \cite{Mermin2007,Jaeger2007}.
\end{comment}
It has been 36 years since Feynman~\cite{Feynman1982Simulating}
proposed the idea about quantum computation. Although there are many
models for quantum computing, the circuit model is still one of the
most widespread models~\cite{544199} no matter whether in theoretical
study or in building a real quantum computation. Information in a
quantum circuit is stored in qubits. Each qubit is a two-dimensional
quantum system. For a two-qubit system, if the first and second qubit
are described as the state~$\ket{\psi}=\alpha_{0}\ket{0}+\alpha_{1}\ket{1}$
and $\ket{\phi}=\beta_{0}\ket{0}+\beta_{1}\ket{1}$, the state of
the whole system is their tensor product 
\begin{equation}
\begin{aligned} & \ket{\psi}\ket{\phi}=\ket{\psi}\otimes\ket{\phi}=\left(\alpha_{0}\ket{0}+\alpha_{1}\ket{1}\right)\otimes\left(\beta_{0}\ket{0}+\beta_{1}\ket{1}\right)\\
= & \alpha_{0}\beta_{0}\ket{0}\otimes\ket{0}+\alpha_{0}\beta_{1}\ket{0}\otimes\ket{1}+\alpha_{1}\beta_{0}\ket{1}\otimes\ket{0}+\alpha_{1}\beta_{1}\ket{1}\otimes\ket{1}\\
= & \alpha_{0}\beta_{0}\ket{0}\ket{0}+\alpha_{0}\beta_{1}\ket{0}\ket{1}+\alpha_{1}\beta_{0}\ket{1}\ket{0}+\alpha_{1}\beta_{1}\ket{1}\ket{1}
\end{aligned}
\end{equation}
.

>>>>

Given a normalized pure state $\ket{\Psi}=\sum_{i=0}^{D-1}\alpha_{i}\ket{i}$,
a natural quantity characterizing an $D$-dimensional system is its
\textit{density matrix}, $\rho=\op{\Psi}{\Psi}$, or 
\begin{equation}
\rho=\begin{bmatrix}|\alpha_{0}|^{2} & \alpha_{0}\alpha_{1}^{*} & \cdots & \alpha_{0}\alpha_{D-1}^{*}\\
\alpha_{1}\alpha_{0}^{*} & |\alpha_{1}|^{2} & \cdots & \alpha_{1}\alpha_{D-1}^{*}\\
\vdots & \vdots & \ddots & \vdots\\
\alpha_{D-1}\alpha_{0}^{*} & \cdots & \alpha_{D-1}\alpha_{D-2}^{*} & |\alpha_{D-1}|^{2}
\end{bmatrix}\,.\label{nqDensity.eq}
\end{equation}
We can now use the complex generalization of the classical Veronese
coordinate system for projective geometry to remove the overall phase
ambiguity $\rme^{\rmi\theta}$ from the $D$-dimensional states. If
we take a particular weighting of the elements of the density matrix~$\rho$,
we can construct a \textit{unit vector} of real dimension $D^{2}$
with the form: 
\begin{equation}
\Hat{a}=\left(|\alpha_{i}|^{2},\ldots,\sqrt{2}\,\mathrm{Re}\ \alpha_{i}\alpha_{j}^{*},\ldots,\sqrt{2}\,\mathrm{Im}\ \alpha_{i}\alpha_{j}^{*},\ldots\right)\,,\label{nqCoords.eq}
\end{equation}
where 
\begin{equation}
\Hat{a}\cdot\Hat{a}=\sum_{i=0}^{D-1}\left(\left|\alpha_{i}\right|^{2}\right)^{2}+\sum_{i=0}^{D-1}\sum_{\substack{j=0\\
j\ne i
}
}^{D-1}\left(\mathrm{Re}\ \alpha_{i}\alpha_{j}^{*}\right)^{2}+\left(\mathrm{Im}\ \alpha_{i}\alpha_{j}^{*}\right)^{2}=\left(\sum_{i=0}^{D-1}\left|\alpha_{i}\right|^{2}\right)\left(\sum_{j=0}^{D-1}\left|\alpha_{j}\right|^{2}\right)=1\,.
\end{equation}
This construction gives an explicit embedding of the $\left(D-1\right)$-dimensional
complex, or $\left(2D-2\right)$-dimensional real, object in a real
space of dimension $D^{2}$. However, this is somewhat subtle because
the vector is of unit length, so technically the embedding space is
a sphere of dimension $D^{2}-1$ embedded in $\R^{D^{2}}$. For example,
the two-dimensional irreducible states could be represented in a four-dimensional
embedding, but the magnitude of every coordinate would be one; furthermore,
the object embedded in the resulting $\Sphere{3}$ is indeed $\Sphere{2}$
because we can fix one complex coordinate to be unity, and let one
vary, giving a total of two irreducible dimensions. In fact one must
choose \textit{two} coordinate patches, one covering one pole of $\Sphere{2}$
with coordinates $\alpha_{0}=0$ and $\alpha_{1}=x_{1}+\rmi y_{1}$,
and the other patch covering the other pole of $\Sphere{2}$ with
coordinates $\alpha_{0}=x_{0}+\rmi y_{0}$ and $\alpha_{1}=0$.

@@

\begin{comment}
An $n$-qubit pure state in the conventional quantum theory (CQT)
is represented as a vector in the $2^{n}$-dimensional Hilbert space.
If we eliminate all symmetries, an irreducible state is actually a
point in the projective Hilbert space \cite{MosseriDandoloff2001,Jaeger2007}
also known as the complex projective space~$\CP{2^{n}-1}$ \cite{Hatcher2001,Bengtsson2007}.
These irreducible quantum states can also be classified as product
states and entangled states, and the latter one plays the essential
role for pure-state algorithms \cite{Mermin2007,Jaeger2007}.
\end{comment}
It has been 36 years since Feynman~\cite{Feynman1982Simulating}
proposed the idea about quantum computation. Although there are many
models for quantum computing, the circuit model is still one of the
most widespread models~\cite{544199} no matter whether in theoretical
study or in building a real quantum computation. Information in a
quantum circuit is stored in qubits. Each qubit is a two-dimensional
quantum system. For a two-qubit system, if the first and second qubit
are described as the state~$\ket{\psi}=\alpha_{0}\ket{0}+\alpha_{1}\ket{1}$
and $\ket{\phi}=\beta_{0}\ket{0}+\beta_{1}\ket{1}$, the state of
the whole system is their tensor product 
\begin{equation}
\begin{aligned} & \ket{\psi}\ket{\phi}=\ket{\psi}\otimes\ket{\phi}=\left(\alpha_{0}\ket{0}+\alpha_{1}\ket{1}\right)\otimes\left(\beta_{0}\ket{0}+\beta_{1}\ket{1}\right)\\
={} & \alpha_{0}\beta_{0}\ket{0}\otimes\ket{0}+\alpha_{0}\beta_{1}\ket{0}\otimes\ket{1}+\alpha_{1}\beta_{0}\ket{1}\otimes\ket{0}+\alpha_{1}\beta_{1}\ket{1}\otimes\ket{1}\\
={} & \alpha_{0}\beta_{0}\ket{0}\ket{0}+\alpha_{0}\beta_{1}\ket{0}\ket{1}+\alpha_{1}\beta_{0}\ket{1}\ket{0}+\alpha_{1}\beta_{1}\ket{1}\ket{1}
\end{aligned}
\end{equation}
in the four-dimensional Hilbert space. For a $n$-qubit system, if
we describe the $j$-th qubit as the state~$\ket{\psi_{j}}$, the
state of the whole system is tensor product of all subsystems $\ket{\Psi}=\ket{\psi_{1}}\otimes\cdots\otimes\ket{\psi_{j}}\otimes\cdots\otimes\ket{\psi_{n}}$
in the Hilbert space of dimension $2^{n}$.
